---
title: "analysis"
---

# Sample analysis

## Load the packages

With this chunk of code we load the packages needed for the analysis.

```{r}
#| output: false
library(babynames)
library(knitr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(pheatmap)
```

Load the custom functions. Use the `#| code-fold: true` to hide the code

```{r}
#| code-fold: true
#
get_most_frequent <- function(babynames, select_sex, from = 1950) {
  most_freq <- babynames |>
    filter(sex == select_sex, year > from) |>
    group_by(name) |>
    summarise(average = mean(prop)) |>
    arrange(desc(average))
    
  return(list(
    babynames = babynames,
    most_frequent = most_freq,
    sex = select_sex,
    from = from))
}

plot_top <- function(x, top = 10) {
  topx <- x$most_frequent$name[1:top]
  
  p <- x$babynames |>
    filter(name %in% topx, sex == x$sex, year > x$from) |>
    ggplot(aes(x = year, y = prop, color = name)) +
    geom_line() +
    scale_color_brewer(palette = "Paired") +
    theme_classic()
  
  return(p)
}
```

## Processing

In this chunk of code we print the head of the table

```{r}
head(babynames)
```

Here we also render the table

```{r}
head(babynames) |> kable()
```

Try to access a file outside the main directory

```{r}
dir("../test/scr/")
```

## Plotting

Plotting them for girls like this:

```{r}
#| label: fig-line-girls
#| fig-cap: "Line plot displaying proportion of top 10 girl names by year"
get_most_frequent(babynames, select_sex = "F") |> plot_top()
```

Plotting them for boys like this:

```{r}
#| label: plot-line-boys
#| fig-cap: "Line plot displaying proportion of top 10 boy names by year"
get_most_frequent(babynames, select_sex = "M") |> plot_top()
```
